---
title: "Analyzing Super Bowl Ads"
description: |
  Like millions of viewers who tune into the big game year after year, I love Super Bowl commercials. I love them so much, in fact, that I wanted to know everything about them… by analyzing and categorizing them, of course. I dug into the defining characteristics of a Super Bowl ad, then grouped commercials based on which criteria they shared — and let me tell you, I found some really interesting clusters of commercials.
editor_options: 
  chunk_output_type: console
---

# Importing Libraries 

```{r setup, include=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(rmarkdown)

knitr::opts_chunk$set(echo = TRUE)
```

# Load Data

```{r Load}
youtube <- read.csv("youtube.csv")
```

# Exploration 

Let's take a look into our data set, and the first thing that I will be doing here, will be exploring the data set, trying to find patterns and understand how the data is distributed.

## Which brand invested the most in advertisements? 

```{r}
youtube %>% 
  count(brand, sort = TRUE) %>% 
  head(20) %>% 
  mutate(brand=fct_reorder(brand,n)) %>% 
  ggplot(aes(n,brand,fill=brand)) +
  geom_bar(stat="identity") +
  theme(legend.position = "none") + 
  labs(title = "Number of Ads per Brand",
       y = "Brand",
       x= "# Of Ads")
```

- We can see that the "beer" companies invested the most: Bud Light and Budweiser

## How is the number of advertisements over the years by brands? 

```{r}
youtube %>% 
  ggplot(aes(year,fill=brand)) +
  geom_bar() + 
  facet_wrap(~ brand) + 
  theme(legend.position = "none")
```

- We can see that Bud Light have decreased the number of advertisements and also in the order hand, KIA stayed very consistent over the years. 

## Number of Views 

Since we are dealing with huge numbers, we will need a log scale for that:

```{r}
youtube %>% 
  gather(metric, value, contains("_count")) %>% 
  ggplot(aes(value)) +
  geom_histogram(binwidth = .5) + 
  scale_x_log10(labels=comma) + 
  labs(x= "# Of Views") + 
  facet_wrap(~ metric)
  
```

## Number of View Counts per Brand 

```{r}
youtube %>%  
  filter(!(is.na(view_count))) %>% 
  mutate(brand=fct_reorder(brand, view_count)) %>% 
  ggplot(aes(view_count, brand)) + 
  geom_boxplot() +
  scale_x_log10()
```

- We can see that NFL and Doritos have the highest number of views, while KIA and Hyundai have the lowest ones. 


## Funny Commercials have Impact of the Number of Views? 

```{r}
youtube %>%  
  filter(!(is.na(view_count))) %>% 
  mutate(brand=fct_reorder(brand, view_count)) %>% 
  ggplot(aes(view_count, brand, fill=funny)) + 
  geom_boxplot() +
  scale_x_log10(labels = comma)
```

- We can not see by the plot above, a direct relatioships between funny ads and number of views. But we can say that the KIA nad Hyundai funny ads are more viewed. 

## Analyzing the Number of Views per Year

```{r}
youtube %>%  
  ggplot(aes(year, view_count, group= year)) + 
  geom_boxplot() +
  scale_y_log10(labels=comma)
```

- Looking on the plot above we can not see clearly the trend of the number of views throughout the years. Perhaps we can have a better view looking into the median of views: 


## Median of Number of View per Year 

```{r}
youtube %>% 
  filter(!is.na(view_count)) %>% 
  group_by(year) %>% 
  summarize(n=n(),
            median_views = median(view_count)) %>%
  mutate(n>=5) %>% 
  ggplot(aes(year,median_views)) +
  geom_line() +
  geom_point(aes(size=n)) +
  theme(legend.position="none") + 
  labs(y="Median # of Views of Super Bowl Ads")
```

- Let's take a look more deeply on the super high median number in 2017: 

```{r}
filtered <- youtube %>% 
  filter(year==2017) %>% 
  arrange(desc(view_count)) 
paged_table(filtered, options = list(rows.print = 10, cols.print = 5))
```

- We can see that the ads with the most number of views was from Budweiser, let's watch the video of the remarcable ads: 

<iframe width="560" height="315" src="https://www.youtube.com/embed/HtBZvl7dIu4" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe> 

## Invetigating the category of Ads

```{r}
youtube %>% 
  filter(!is.na(view_count)) %>%
  gather(category, value, funny:use_sex) %>% 
  group_by(category,value) %>% 
  summarize(n= n(),
             median_view_count=median(view_count)) %>% 
  ggplot(aes(category,median_view_count,fill=value))+
  geom_col(position = "dodge")
```

- Showing a little bit of how significant patriotism is for Americans, we can see that add which includes patriotic in it, has a significant higher number of views. The same happen with danger, but to total opposite happen with the use of sex. 

## What type of Ads is more comoom over time? 

```{r}
youtube %>% 
  gather(category, value, funny:use_sex) %>% 
  group_by(category=str_to_title(str_replace_all(category, "_"," ")),
           year) %>% 
  summarize(pct=mean(value),
            n = n()) %>% 
  ggplot(aes(year,pct,color=category))+
  geom_line()+
  scale_y_continuous(labels=percent)+
  facet_wrap(~ category) +
  theme(legend.position = "none") + 
  labs(x = "Time (rounded to 2-years)",
       y = "% of ads with the quality ")
```

- Now we can identify some trends: 
  - The used of Celebrity has been increasing over last 5 years; 
  - And the ads with Sex content has been decreasing over time; 
  - Even with all the critical even that happened in 2009, he patriotic ads didn't increasing after it. 


## Binomial Regression

But to be 100% sure of this, we can run a regression and look over the p values: 

```{r}
# For Use Sex
glm (use_sex ~ year, 
     data = youtube,
     family = "binomial") %>% 
  summary()

# For Celebrity
glm (celebrity ~ year, 
     data = youtube,
     family = "binomial") %>% 
  summary()
```

- Now, we can see even more clearly that the "use_sex" is more significant than "celebrity", but with opposite signs. 

```{r}
gathered_categories <- youtube %>% 
  gather(category, value, funny:use_sex) %>% 
  mutate(category=str_to_title(str_replace_all(category, "_"," ")))
```

## Type of Ads for each company thru Years


```{r}
library(tidytext)
gathered_categories %>% 
  group_by(brand, category) %>% 
  summarize( pct = mean(value)) %>% 
  ungroup() %>% 
  mutate(brand = reorder_within(brand, pct, category)) %>% 
  ggplot(aes(pct, brand))+
  geom_col()+
  scale_x_continuous(labels = percent)+
  scale_y_reordered()+
  facet_wrap(~ category, scales = "free_y") +
  labs( y = "",
        x= "% of brands ads have this quality")
  
```

- We can see that:
  - Coca Cola is the one who uses the most Animals in their Ads, and NFL never uses; 
  - NFL shows almost 75% at the time, celebrities in their ads;
  - Doritos uses funny ads in almost 100% at the time. 
  
  
## Difference between types of ads in Brands 
  
```{r, fig.width=11, fig.height=8, dpi=300}
gathered_categories %>% 
  group_by(brand, category) %>% 
  summarize( pct = mean(value)) %>% 
  ungroup() %>% 
  mutate(category = reorder_within(category, pct, brand)) %>% 
  ggplot(aes(pct, category))+
  geom_col()+
  scale_x_continuous(labels = percent)+
  scale_y_reordered()+
  facet_wrap(~ brand, scales = "free_y") +
  labs( y = "",
        x= "% of brands ads have this quality",
        title = "What is each brand's fingerprint in terms of the type of ads it produces?")
```

- Let's imagine that someone came to you with this question: What is the difference between Coca Cola and Pepsi in terms of their SuPer Bowl ads? 
  - Their both tries to show their products quickly, where Pepsi uses more Funny ads and Coca Cola uses Animals. 
  - Pepsi often uses Sex and Coca Cola never uses. 
  
  
```{r}
by_brand_categories <- gathered_categories %>% 
  group_by(brand, category) %>% 
  summarize( pct = mean(value)) %>% 
  ungroup() 
```
  
  
## Heat Map

```{r}
by_brand_categories %>% 
  reshape2::acast(brand~category, value.var = "pct") %>% 
  heatmap()
```

- Here we can see some correlations between brands and categories:
  - Pepsi and KIA seems to have the same type of ads;
  - Hyundai, Bud Light and Doritos uses a lot of funny ads;
  - Coca Cola and Pepsi seems to have the same type of ads too. 
  
  
# Likes and Dislikes 

Let's take a look on the ratio of dislikes and likes for those brands: 

```{r}
likes_dislikes <- youtube %>% 
  mutate(dislike_pct = dislike_count/view_count,
         like_ratio = like_count/view_count,
         like_dislike_total = like_count + dislike_count,
         dislike_pct = dislike_count/(like_count + dislike_count)) %>% 
  filter(like_dislike_total >= 1000) %>% 
  filter(!is.na(dislike_pct)) %>% 
  select(brand, year, description, view_count, like_dislike_total, 
         like_count, dislike_count, like_ratio) %>% 
  arrange(desc(view_count))
``` 

```{r}
likes_dislikes %>% 
  mutate(dislike_pct = dislike_count/(like_count + dislike_count),
         brand = fct_reorder(brand, dislike_pct, mean)) %>% 
  ggplot(aes(dislike_pct, brand)) +
  geom_boxplot() + 
  scale_x_continuous(labels = percent) +
  labs(title="What brands tend to produce polarizing ads in terms of Youtube likes?",
       x = "Dislike / (Likes + Dislikes)",
       y = "Brand")
```

- Finally here, we see that Coca Cola has never produced an ad that got less than 4% of dislike percentage and the mean of about 16%. 

# End 

We can do way more more where, but I will stay with those analysis for now. 
Thank you! 

  