---
title: "Data Cleaning in SQL"
description: |
  Continuing my SQL projects, I will be performing data cleaning in a more advanced way. The data set that I will be cleaning have the information about Nashville Housing, that contains information about the Property Address, Sales Date, Sales Price, Owner Name and much more.
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing Libraries 

```{r Connections}
library(RODBC)
library(DBI)
library(odbc)
```

# Establishing Connection

In order to use SQL databases, you must establish connection with Microsoft SQL Server. One way to do that, is using "dbConnect":

```{r}
conn2 <- dbConnect(odbc:: odbc(),"SQLServer_DSN" )
```

# Cleaning Data

## First View

Let's take a first look into our data:

```{sql connection=conn2, output.var="firstview"}
SELECT * 
FROM "Portfolio"."dbo"."nashville"
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(firstview, options = list(rows.print = 10, cols.print = 19))
```

## Changing Data Format

```{sql eval=FALSE, connection=conn2, include=TRUE}
ALTER TABLE Portfolio.dbo.nashville 
ADD SaleDateConverted Date; 

UPDATE Portfolio.dbo.nashville 
SET SaleDateConverted = CONVERT(Date,SaleDate)
```

```{sql connection=conn2, output.var="selectdata"}
SELECT SaleDateConverted, SaleDate	
FROM Portfolio.dbo.nashville
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(selectdata, options = list(rows.print = 10, cols.print = 5))
```

- Now this new column is in "date" format


## Populate Property Address data

First let's take a look if there is missing values on the PropertyAddress column. 

```{sql connection=conn2}
SELECT PropertyAddress
FROM "Portfolio"."dbo"."nashville"
WHERE PropertyAddress IS NULL
```

- We can see that we do have missing values in this column, but how we will populate those? Based on what? 
  - Let's take a look on our other column, ParcelID: 

```{sql connection=conn2, output.var="lookdata"}
SELECT ParcelID, PropertyAddress
FROM "Portfolio"."dbo"."nashville"
ORDER BY ParcelID
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(lookdata, options = list(rows.print = 10, cols.print = 5))
```

- We can see that the **ParcelID** *015 14 0 060.00* on page 5, have the same **PropertyAddress**. This means that we can populate our **PropertyAddress** column looking into the **ParcelID** values, if there is a missing value on that column, we will check on the other one to see if we can find the respective address. 
  - In order to do that, we will need a **self join**: 

```{sql connection=conn2, output.var="infecdata"}
Select a.ParcelID, a.PropertyAddress, b.ParcelID, b.PropertyAddress, ISNULL(a.PropertyAddress,b.PropertyAddress) AS NewPropertyAddress
From "Portfolio"."dbo"."nashville" a
JOIN "Portfolio"."dbo"."nashville" b
	on a.ParcelID = b.ParcelID
	AND a.[UniqueID ] <> b.[UniqueID ]
Where a.PropertyAddress is null
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(infecdata, options = list(rows.print = 10, cols.print = 5))
```

- Great, we can see our new column that will be populate the missing values. Let's update: 

```{sql connection=conn2}
Update a
SET PropertyAddress = ISNULL(a.PropertyAddress,b.PropertyAddress)
From "Portfolio"."dbo"."nashville" a
JOIN "Portfolio"."dbo"."nashville" b
	on a.ParcelID = b.ParcelID
	AND a.[UniqueID ] <> b.[UniqueID ]
Where a.PropertyAddress is null
```

- Worked perfectly! To be 100% that there is no more missing values in this column, let's doble check: 

```{sql connection=conn2}
Select a.ParcelID, a.PropertyAddress, b.ParcelID, b.PropertyAddress, ISNULL(a.PropertyAddress,b.PropertyAddress) AS NewPropertyAddress
From "Portfolio"."dbo"."nashville" a
JOIN "Portfolio"."dbo"."nashville" b
	on a.ParcelID = b.ParcelID
	AND a.[UniqueID ] <> b.[UniqueID ]
Where a.PropertyAddress is null
```

- 0 rows! Which means that we did it!

## Breaking out Address into Individual Columns (Address, City, State)

In our column, PropertyAddress, we can see that there a comma, separating the address from the city. Let's create a new column to get the name of the cities. 

- First, the Address:

```{sql eval=FALSE, connection=conn2, include=TRUE}
ALTER TABLE Portfolio.dbo.nashville
ADD PropertySplitAddress Nvarchar(255);

Update Portfolio.dbo.nashville
SET PropertySplitAddress = SUBSTRING(PropertyAddress, 1, CHARINDEX(',', PropertyAddress) -1 )
```

- Second, the City:

```{sql eval=FALSE, connection=conn2, include=TRUE}
ALTER TABLE Portfolio.dbo.nashville
ADD PropertySplitCity Nvarchar(255);

Update Portfolio.dbo.nashville
SET PropertySplitCity = SUBSTRING(PropertyAddress, CHARINDEX(',', PropertyAddress) + 1 , LEN(PropertyAddress))
```


```{sql connection=conn2, output.var="contryhigh" }
Select PropertySplitAddress, PropertySplitCity
From Portfolio.dbo.nashville
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown) 
paged_table(contryhigh, options = list(rows.print = 10, cols.print = 2))
```

- We can see that both new columns were added *PropertySplitAddress* and *PropertySplitCity*. 

We also can do something very similar to get information on Address, City and State for the *OwnerAddress* column. But first, let's look on this column:

```{sql connection=conn2}
SELECT OwnerAddress
FROM "Portfolio"."dbo"."nashville"
```

- I will use the function `PARSENAME` to get this information: 

```{sql eval=FALSE, connection=conn2, include=TRUE}
-- Address
ALTER TABLE Portfolio.dbo.nashville
Add OwnerSplitAddress Nvarchar(255);

Update Portfolio.dbo.nashville
SET OwnerSplitAddress = PARSENAME(REPLACE(OwnerAddress, ',', '.') , 3)
```


```{sql eval=FALSE, connection=conn2, include=TRUE}
-- City
ALTER TABLE Portfolio.dbo.nashville
Add OwnerSplitCity Nvarchar(255);

Update Portfolio.dbo.nashville
SET OwnerSplitCity = PARSENAME(REPLACE(OwnerAddress, ',', '.') , 2)
```



```{sql eval=FALSE, connection=conn2, include=TRUE}
-- State
ALTER TABLE Portfolio.dbo.nashville
Add OwnerSplitState Nvarchar(255);

Update Portfolio.dbo.nashville
SET OwnerSplitState = PARSENAME(REPLACE(OwnerAddress, ',', '.') , 1)
```


```{sql connection=conn2, output.var="contryhigh"}
SELECT OwnerSplitAddress, OwnerSplitCity, OwnerSplitState
FROM Portfolio.dbo.Nashville
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(contryhigh, options = list(rows.print = 10, cols.print = 5))
```

- We can see that three new column were successfully created, with the information that we want to. 

## Change Y and N to Yes and No in "Sold as Vacant" field

Let's change the values on *SoldAsVacant* column: Y -> Yes and N -> No. 

```{sql connection=conn2}
SELECT DISTINCT(SoldAsVacant), Count(SoldAsVacant)
FROM "Portfolio"."dbo"."nashville"
Group by SoldAsVacant
order by 2
```

```{sql connection=conn2}
Update Portfolio.dbo.nashville
SET SoldAsVacant = CASE When SoldAsVacant = 'Y' THEN 'Yes'
	   When SoldAsVacant = 'N' THEN 'No'
	   ELSE SoldAsVacant
	   END
```

- Now, let's see the results:

```{sql connection=conn2}
SELECT DISTINCT(SoldAsVacant), Count(SoldAsVacant)
FROM "Portfolio"."dbo"."nashville"
Group by SoldAsVacant
order by 2
```


## Remove Duplicates

Let's find out all the duplicate values in our data set:

```{sql connection=conn2, output.var="contdeath"}
WITH RowNumCTE AS(
Select *,
	ROW_NUMBER() OVER (
	PARTITION BY ParcelID,
				 PropertyAddress,
				 SalePrice,
				 SaleDate,
				 LegalReference
				 ORDER BY
					UniqueID
					) row_num

FROM "Portfolio"."dbo"."nashville"
)

Select *
From RowNumCTE
Where row_num > 1
Order by PropertyAddress
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(contdeath, options = list(rows.print = 10, cols.print = 30))
```

And delete them:

```{sql connection=conn2}
WITH RowNumCTE AS(
Select *,
	ROW_NUMBER() OVER (
	PARTITION BY ParcelID,
				 PropertyAddress,
				 SalePrice,
				 SaleDate,
				 LegalReference
				 ORDER BY
					UniqueID
					) row_num

FROM "Portfolio"."dbo"."nashville"
)

DELETE
From RowNumCTE
Where row_num > 1
```

- Let's check if this worked:

```{sql connection=conn2, output.var="contdeath1"}
WITH RowNumCTE AS(
Select *,
	ROW_NUMBER() OVER (
	PARTITION BY ParcelID,
				 PropertyAddress,
				 SalePrice,
				 SaleDate,
				 LegalReference
				 ORDER BY
					UniqueID
					) row_num

FROM "Portfolio"."dbo"."nashville"
)

Select *
From RowNumCTE
Where row_num > 1
Order by PropertyAddress
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(rmarkdown)
paged_table(contdeath1, options = list(rows.print = 10, cols.print = 30))
```
- Perfect, there is no more duplicate values!


## Delete Unused Colunms

```{sql eval=FALSE, connection=conn2, include=TRUE}
ALTER TABLE Portfolio.dbo.nashville
DROP COLUMN OwnerAddress, TaxDistrict, PropertyAddress, SaleDate;
```

